Lattice Boltzmann study of mass transfer for complicated geometries

### 2012-01-27 Text/introduction.tex
Writing a report.

\newcommand{\beq}{\begin{equation}}
\newcommand{\feq}{\end{equation}}
\newcommand{\beqstar}{\begin{equation*}}
\newcommand{\feqstar}{\end{equation*}}                                     
\newcommand{\beqal}{\begin{equation}\begin{aligned}}
\newcommand{\feqal}{\end{aligned}\end{equation}}
\newcommand{\beqalstar}{\begin{equation*}\begin{aligned}}
\newcommand{\feqalstar}{\end{aligned}\end{equation*}}
\newcommand{\vol}{k_L\,a}
\newcommand{\lbubble}{L_{\mathrm{bubble}}}
\newcommand{\lunit}{L_{\mathrm{unit}}}
\newcommand{\lslug}{L_{\mathrm{slug}}}
\newcommand{\ububble}{U_{\mathrm{bubble}}}
\newcommand{\uliq}{U_{\mathrm{liq}}}
\newcommand{\ugas}{U_{\mathrm{gas}}}
\newcommand{\uoutlet}{U_{\mathrm{outlet}}}
\newcommand{\cbubble}{C_{\mathrm{bubble}}}
\newcommand{\cinlet}{C_{\mathrm{inlet}}}
\newcommand{\coutlet}{C_{\mathrm{outlet}}}

@@

Other correlations can be found in works of \citet{kreutzer-overview}
and \citet{bercic-mass}. Among a variety of experimental works extensively covered
\cite{yue-mass}, the numerical validations to the best authors' knowledge are sparse. One of the
reasons is that the experimental correlations are established for continous system, i.e. one does
not distinguish separate bubbles, but looks to the system as a whole measuring mass concentration
in different places. Thus, numerical simulations require a significant computational
resources to be able to simulate enough number of the unit cells to mimic the continuous picture.
There are some works as how to avoid large scale numerical simulations by performing numerical
simulations for one unit cell. For example, \citet{vanbaten-circular} perform numerical
simulations for one unit cell using periodic boundary conditions and measure the volumetric mass
transfer coefficient as:
\begin{equation}
\label{main:simulation:equation}
k_L a=\frac{\mathrm{Flux}}{\cbubble-\coutlet} \frac{\mathrm{bubble\ surface\ area}}{\mathrm{unit\
cell\ volume}},
\end{equation}
where $\coutlet=\int{C \uoutlet \mathrm{d}A}/\int{\uoutlet\mathrm{d}A}$. 
The concentration flux is calculated as the difference between overall
average concentration in the whole domain ($C_{overall}=\int_{V} C \mathrm{d}V /V$)
at time
$t_1$ and at time $t_2$ divided on the time difference $t_2-t_1$. The agreement between numerical
simulations and correlations of \citet{bercic-mass} was good. However, the numerical model only
accounts for circular capillaries and short contact times when the
bubble film is not saturated with tracer. As well, authors took the
approximation of bubble shapes for small capillary numbers $Ca<0.01$ and then derived the
hydrodynamics fields by imposing a free-slip condition at the bubble surface. {\color{red} Give
another numerical work by Kreutzer.}

@@

There are different methods to estimate the mass transfer coefficient $k_L$. We first examine the
theoretical definitions of the mass transfer in case of point mass sources.
\subsection{Point mass sources}
In what follows we will present three approaches to calculate point mass transfer
coefficients:
\begin{enumerate}
\item
Let us look at the infinitesimal small domain of the volume $A \Delta x$ not
moving and with the point mass source. Then the concentration difference can be found as $\Delta C = C^* -
C(t)$, where $C^*$ is the imposed point source concentration, $C(t)$ is
the time dependent average concentration. Therefore one can write the time dependent ODE for the
average concentration in the domain:
\beq
\dot{m}= A \Delta x \frac{\mathrm{d}C}{\mathrm{d} t} = k_L P (C^{*}-C(t)), 
\feq
with the initial condition $C(0)=0$
The solution can be found by solving ODE:
\beq
C(t)= C^{*}(1-\exp(-\vol t )), 
\feq
where $\vol$ is the volumetric mass transfer coefficient defined as:
\beq
\vol=k_L \frac{P}{V},
\feq
where $P$ is the source surface, $V$ is the unit cell volume.
\item
The previous consideration is able to predict mass transfer happening in moving with the velocity
$U$ liquid, see Fig. \ref{fig:moving:frame}. 
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/mass_transfer.eps}
\caption{The mass transfer in the moving liquid. \label{fig:moving:frame}}
\end{figure}

If one can assume that the mass point sources are distributed in the whole medium, then the mass accumulated in the unit volume $V=A \Delta x$ can be calculated as the difference of 
mass fluxes entering and leaving domain $U \bigl(C(x+\Delta x)-C(x)\bigr)$. We assume that the
source is a point source with the effective perimeter (area) $P$. The accumulated mass should be proportional
to the mass transfer coefficient:
\beq
U \bigl(C(x+\Delta x)-C(x)\bigr)=k_L P \bigl(C^{*}-C(x)\bigr), 
\feq 
giving the same equation but only in the coordinate domain:
\beq
C(x)= C^{*}\Bigl(1-\exp\bigl(-k_L a \frac{x}{U} \bigr)\Bigr).
\label{main:mass:transfer:expression} 
\feq

@@

The expression (\ref{main:mass:transfer:expression}) was used in the experiments by
\citet{bercic-mass}. One should be accurate with the definition of velocities because we have two
different phases. Usually, one can take the velocity $U$ to be as
average velocity or $U=U_{gas}+U_{liq}$, where $U_{gas}$ and $U_{liq}$ are liquid and gas
superficial velocities correspondingly. In this case, the continuouty assumption is restored and
all considerations above are valid.

 However, if one wants to analytically estimate the mass transfer coefficients, the situation is
much more complicated because of the two phases presence. Phases can have different mixing patterns
from film to slug and from slug to film. The analytical approaches described above assume that the
contributions from film and bubble caps can be calculated separately. However one can see that such 
an assumption for bubble caps (slug well mixed and the concentration is uniformly distributed over
the slug) usually overpredicts the mass transfer \cite{irandoust}. This happens since some tracer
concentration from film is mixed with the slug and changes the overall concentration in the slug.
Another issue is that if the slug is long enough then the film saturates with the tracer
concentration and its influence on the mass transfer can be negligible \cite{vanbaten-circular}.
Mixing patterns of the film and the liquid slugs are of great importance for the analytical
estimation of the mass transfer \cite{yue-mass}.

\subsection{Numerical calculation}

@@

The numerical  procedure of studying mass transfer is presented in the following steps:
\begin{description}
 \item[Flow field] The hydrodynamics for the bubble motion in the periodic domain is obtained
according to the work \cite{kuzmin-binary2d}. The grid resolution is taken to insure grid
independency of results.  
 \item[Bubble reference frame] Once the hydrodynamics is resolved, the mass transfer simulations
are conducted in the moving
reference frame with the bubble. Then the bubble stands still and the flow is coming around the
bubble. We impose a steady concentration of solutant on the surface of the bubble.
\item[Mass transfer] One needs not only perform a Galilean transformation to the reference
frame moving with the bubble but also to
transfer the bubble to the inlet condition. This can insure that the studied phenomenon is a liquid
slug. 
\end{description}

In the numerical simulation it is not possible to simulate really long domains with the infinite
number of the bubbles. As we saw before there are two approaches towards it - either to simulate
the bubble train and then to measure concentration along the pipe, Eq.
\ref{main:mass:transfer:expression}, or to transfer to the reference frame moving with bulk
velocity $U$ and conduct the same measurements through Eq. \ref{main:mass:transfer:expression}.
However, both methods do require a tracking of moving bubbles which is complicated from the
numerical point of view.

Therefore, one needs to come up with the simple smaller domain calculations
of the mass transfer coefficient, which represent the continuous picture with infinite numbers of
the bubbles closely. In comparison with analytical estimations of the mass transfer coefficient,
where the mixing patterns are extremely important, the numerical simulations need only to mimic
the continous picture and provide the way to calculate the volumetric mass transfer coefficient.

We perform simulations in the frame moving with the bubble (velocity
$U_{bubble}$), that the bubble position stands steady. The bubble velocity $U_{bubble}$ is
different from the bulk velocity $U_{gas}+U_{liq}$. In this case we need to perform coordinate
change:
\beqal
&x(t)=U_{bubble}t\\
&C(x)=C^{*}\Bigl(1-\exp\bigl(-k_L\, a \frac{x}{U_{gas}+U_{liq}}\bigr)\Bigr)\\
&C(t)=C^{*}\Bigl(1-\exp\bigl(-k_L\, a\, t \frac{U_{bubble}}{U_{gas}+U_{liq}}\bigr)\Bigr)
\feqal
If the unit volume that we simulate is quite long one can obtain the values of the concentration
in the inlet and outlet as a function of time:
\beqal
&C(U_{bubble}\, t)=C^{*}\Bigl(1-\exp\bigr(-k_L\, a\, t
\frac{U_{bubble}}{U_{gas}+U_{liq}}\bigr)\Bigr)\\
&C(U_{bubble}\,t+L_{unit})=C^{*}\Bigl(1-\exp\bigl(-k_L a \frac{
U_{bubble}\,t+L_{unit}}{U_{gas}+U_{liq}}\bigr)\Bigr)\\
&=C^{*}\Bigl(1-\exp\bigl(-k_L\, a\, t \frac{U_{bubble}}{U_{gas}+U_{liq}}\bigr)\exp\bigl(-k_L a
\frac{L_{unit}}{U_{gas}+U_{liq}}\bigr)\Bigr)
\feqal
Therefore the mass transfer coefficient can be calculated as follows:
\begin{equation}
\label{theor:continuous:mass:transfer}
\begin{aligned}
&-k_L a \frac{L_{unit}}{U_{gas}+U_{liq}}=\ln\Bigl(\frac{C^{*}-C_{outlet}}{C^{*}-C_{inlet}}\Bigr)\\
&k_L a
=\frac{U_{gas}+U_{liq}}{L_{unit}}\ln\Bigl(\frac{C^{*}-C_{inlet}}{C^{*}-C_{outlet}}
\Bigr)
\end{aligned}
\end{equation}
If inlet and outlet concentrations are small in comparison with the saturation concentration $C^{*}$
then the expression for the volumetric mass transfer coefficitent $k_L a$ can be simplified:
\beq
\vol=\frac{\Delta C}{C^{*}} \frac{U_{gas}+U_{liq}}{L_{unit}},
\feq
where $\Delta C=C_{outlet}-C_{inlet}$. If we go to the continuous formulation by changing $L_{unit}$
to be a continuous coordinate $x$, then the following can be obtained \cite{jos-mass}:
\beq
\Delta C = C^{*} \vol \frac{x}{U_{gas}+U_{liq}}
\label{small:conc:mass:estimation}
\feq

We consider the following possibilities of conducting numerical simulations:
\begin{enumerate}
 \item 
The tracer picking up procedure can be introduced
in the manner of the considerations above. We impose the concentration at the entrance to be
constant $C(0,y)=C_{inlet}$, at the outlet we impose the outflow boundary condition $\frac{\partial
C}{\partial x}(L,y)=0$. To be compliant with the considerations above one needs to impose as well
the outflow boundary conditions for the inlet as well. However, without loss of the generality one
can impose $C_{inlet}=0$, since the walls in the reference frame
moving with the bubble are moving from the inlet to the outlet. That means that no tracer
concentration leaves domain from the inlet side, which is true for the capillary numbers larger
than $0.6$.  However, this boundary condition can not be used in the case of small capillary
numbers, since there is a strong circulation in front of the bubble. One can see the sketch of the
approach in Fig. \ref{fig:benchmark:jos}. The longer the numerical domain length the better will be
results. One can estimate the volumetric mass transfer coefficient in this case through the equation
(\ref{small:conc:mass:estimation}) for small concentrations. Note a few considerations:
\begin{itemize}
\item  The considerations above were established for the homogenous velocity field. In the case of
the non-homogenous velocity field one needs to obtain a concentration from the mass flux:
\beq
C_{outlet}=\frac{\int{U(y) C(L_{unit},y) \mathrm{d}y}}{\int{U(y) \mathrm{d}y}}.
\feq
\item Note, that one cannnot use the definition of the mass transfer coefficient as far as it
varies through the bubble due to inhomogenoities of the flow. Therefore, the volumetric mass
transfer coefficient is introduced that takes into the account the bubble geometry.
\end{itemize}
\begin{figure}
\includegraphics[width=\textwidth]{Figures/benchmark_jos.eps}
\caption{The numerical mass transfer benchmark. \label{fig:benchmark:jos}}
\end{figure}

\item Another possibility is to use the periodic boundary conditions as the limiting case of the
continuous formulation. In this case the inlet and outlet concentrations are not much different
($C(x)=C(x+L_{unit})=C(x+\Delta x)\approx C_{media}$, see Fig. \ref{fig:moving:frame}), where
$C_{media}$ is the mass averaged over domain concentration. This formulation requires the periodic
boundary condition $C_{inlet}=C_{outlet}$. By the definition the mass transfer coefficient is
proportional to the gain of the mass in the system divided by the concentration difference:
\beq
j=\frac{\dot{m}}{P \Delta C}=\frac{\dot{m}}{P (C^{*}-C_{media})},
\feq
where mass difference in the domain can be calculated as:
\beq
\dot{m}=\frac{m_2-m_1}{t_2-t_1}=\frac{\int_{C(t_2)\mathrm{d}V}-\int{C(t_1)\mathrm{d}V}}{t_2-t_1}.
\feq
Therefore the assumptions  of this approach is that the difference between inlet and outlet is not
considerably large and the tracer is well mixed in the slug, as one can take the concentration
difference as $\Delta C = C^{*}-C_{media}$. The best case scenario is when $C_{media}$ is close to
zero, in this case we can be sure that the flux to the domain is calculated precisely and periodic
boundary conditions are fulfilled.
\end{enumerate}

@@

\section{Comparison between different approaches for one unit cell}
We consequently will examine different approaches to the mass transfer definition, Fig.
\ref{fig:comparison:one:unit:cell}:
\begin{description}
\item[I] Definition of \citet{vanbaten-circular} with periodic boundary conditions where a bubble
is located near the entrance of the computational domain, and the characteristic concentration is
the inlet/outlet concentration, Fig \ref{fig:benchmark}. 
\item[II] Definition of \citet{vanbaten-circular} with periodic boundary conditions where a bubble
is located near the entrance of the computational domain, and the characteristic concentration is
the average concentration, Fig \ref{fig:benchmark}.
\item[III] Definition of \citet{vanbaten-circular} with periodic boundary conditions where a
bubble is located in the middle of the computational domain, and the characteristic concentration
is the inlet/outlet concentration.
\item[IV] Definition of \citet{vanbaten-circular} with periodic boundary conditions where a bubble
is located in the middle of the computational domain, and the characteristic concentration is the
average concentration.
\item[V] Definition with the concentration jump between inlet and outlet over one unit cell.
\end{description}
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/benchmark_periodic.eps}\\
\includegraphics[width=\textwidth]{Figures/benchmark_middle.eps}\\
\caption{The two-dimensional benchmarks for the  the mass transfer
coefficient (bottom) for the bubble located near the entrance (top) and at the middle of the domain
(bottom). \label{fig:benchmark}}
\end{figure}

>>>>

\newcommand{\beq}{\begin{equation}}
\newcommand{\feq}{\end{equation}}
\newcommand{\beqstar}{\begin{equation*}}
\newcommand{\feqstar}{\end{equation*}}                                     
\newcommand{\beqal}{\begin{equation}\begin{aligned}}
\newcommand{\feqal}{\end{aligned}\end{equation}}
\newcommand{\beqalstar}{\begin{equation*}\begin{aligned}}
\newcommand{\feqalstar}{\end{aligned}\end{equation*}}
\newcommand{\vol}{k_L\,a}
\newcommand{\lbubble}{L_{\mathrm{bubble}}}
\newcommand{\lunit}{L_{\mathrm{unit}}}
\newcommand{\lslug}{L_{\mathrm{slug}}}
\newcommand{\ububble}{U_{\mathrm{bubble}}}
\newcommand{\uliq}{U_{\mathrm{liq}}}
\newcommand{\ugas}{U_{\mathrm{gas}}}
\newcommand{\uoutlet}{U_{\mathrm{outlet}}}
\newcommand{\cbubble}{C_{\mathrm{bubble}}}
\newcommand{\cinlet}{C_{\mathrm{inlet}}}
\newcommand{\coutlet}{C_{\mathrm{outlet}}}
\newcommand{\coverall}{C_{\mathrm{overall}}}
\newcommand{\cstar}{C^{*}}
\newcommand{\cmedium}{C_{\mathrm{medium}}}
\newcommand{\volnondim}{\vol \frac{\lunit}{\ububble+\ugas}}

@@

Other correlations can be found in works of \citet{kreutzer-overview}
and \citet{bercic-mass}. Among a variety of experimental works extensively covered
\cite{yue-mass}, the numerical validations to the best authors' knowledge are sparse. One of the
reasons is that the experimental correlations are established for continous system, i.e. one does
not distinguish separate bubbles, but looks to the system as a whole measuring mass concentration
in different places. Thus, numerical simulations require a significant computational
resources to be able to simulate enough number of the unit cells to mimic the continuous picture.
There are some works as how to avoid large scale numerical simulations by performing numerical
simulations for one unit cell. For example, \citet{vanbaten-circular} perform numerical
simulations for one unit cell using periodic boundary conditions and measure the volumetric mass
transfer coefficient as:
\begin{equation}
\label{main:simulation:equation}
k_L a=\frac{\mathrm{Flux}}{\cbubble-\coutlet} \frac{\mathrm{bubble\ surface\ area}}{\mathrm{unit\
cell\ volume}},
\end{equation}
where $\coutlet=\int{C \uoutlet \mathrm{d}A}/\int{\uoutlet\mathrm{d}A}$. 
The concentration flux is calculated as the difference between overall
average concentration in the whole domain ($\coverall=\int_{V} C \mathrm{d}V /V$)
at time
$t_1$ and at time $t_2$ divided on the time difference $t_2-t_1$. The agreement between numerical
simulations and correlations of \citet{bercic-mass} was good. However, the numerical model only
accounts for circular capillaries and short contact times when the
bubble film is not saturated with tracer. As well, authors took the
approximation of bubble shapes for small capillary numbers $Ca<0.01$ and then derived the
hydrodynamics fields by imposing a free-slip condition at the bubble surface. {\color{red} Give
another numerical work by Kreutzer.}

@@

There are different methods to estimate the mass transfer coefficient $k_L$. We first examine the
theoretical definitions of the mass transfer in case of point mass sources.
\subsection{Point mass sources}
In what follows we will present three approaches to calculate point mass transfer
coefficients:
\begin{enumerate}
\item
Let us look at the infinitesimal small domain of the volume $A \Delta x$ not
moving and with the point mass source. Then the concentration difference can be found as $\Delta C =
\cstar -
C(t)$, where $\cstar$ is the imposed point source concentration, $C(t)$ is
the time dependent average concentration. Therefore one can write the time dependent ODE for the
average concentration in the domain:
\beq
\dot{m}= A \Delta x \frac{\mathrm{d}C}{\mathrm{d} t} = k_L P (\cstar-C(t)), 
\feq
with the initial condition $C(0)=0$
The solution can be found by solving ODE:
\beq
C(t)= \cstar (1-\exp(-\vol t )), 
\feq
where $\vol$ is the volumetric mass transfer coefficient defined as:
\beq
\vol=k_L \frac{P}{V},
\feq
where $P$ is the source surface, $V$ is the unit cell volume.
\item
The previous consideration is able to predict mass transfer happening in moving with the velocity
$U$ liquid, see Fig. \ref{fig:moving:frame}. 
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/mass_transfer.eps}
\caption{The mass transfer in the moving liquid. \label{fig:moving:frame}}
\end{figure}

If one can assume that the mass point sources are distributed in the whole medium, then the mass accumulated in the unit volume $V=A \Delta x$ can be calculated as the difference of 
mass fluxes entering and leaving domain $U \bigl(C(x+\Delta x)-C(x)\bigr)$. We assume that the
source is a point source with the effective perimeter (area) $P$. The accumulated mass should be proportional
to the mass transfer coefficient:
\beq
U \bigl(C(x+\Delta x)-C(x)\bigr)=k_L P \bigl(\cstar-C(x)\bigr), 
\feq 
giving the same equation but only in the coordinate domain:
\beq
C(x)= \cstar \Bigl(1-\exp\bigl(-k_L a \frac{x}{U} \bigr)\Bigr).
\label{main:mass:transfer:expression} 
\feq

@@

\subsection{Bubble train}
In the application to the bubble train flow, if one can think of one bubble to be a point source,
then all considerations above are valid. For example, the expression
(\ref{main:mass:transfer:expression}) was used in the experiments by
\citet{bercic-mass}. However, one should be accurate with the definition of velocities because two
different phases co-exist in the bubble train flow. Usually, one can take the velocity $U$ to be as
a bulk velocity or $U=\ugas+\uliq$, where $\ugas$ and $\uliq$ are liquid and gas
superficial velocities correspondingly.

Though the experimental measurements are straightforward as measuring concentrations at different
locations and calculating the volumetric mass transfer coefficient through the logarithmic
function. However, if one wants to analytically estimate the mass transfer coefficients, the
situation is much more complicated because of two phases presence and complicated bubble
geometry. Phases can have different mixing patterns
from film to slug and from slug to film. As it was mentioned before depending on the capillary
number the mixing velocity pattern is different. Analytical approaches
\cite{irandoust,vanbaten-circular} assume that the
contributions from film and bubble caps can be calculated separately. However, one can see that
such 
an assumption for bubble caps (slug well mixed and the concentration is uniformly distributed over
the slug) usually overpredicts the mass transfer \cite{irandoust}. This happens since some tracer
concentration from film is mixed with the slug and changes the overall concentration in the slug.
Thus, all estimations for the analytical mass as well do not account mutual mass transfer from
neighbouring bubbles.

Another issue is that if the slug is long enough then the film saturates with the tracer
concentration and its influence on the mass transfer can be negligible \cite{vanbaten-circular}.
Mixing patterns of the film and the liquid slugs are of great importance for the analytical
estimation of the mass transfer \cite{yue-mass}. However, the assumptions usually taken for the
mass transfer calculation are small capillary number and certain mixing patterns which help to
estimate the mass transfer using the penetration theory of \citet{higbie}.

The numerical approach can take into the account the complicated mixing patterns and geometries.
However, there is a  challenge to restore the continous approach as it is seen in experiments. Next
section gives more details about numerical simulations.

\subsection{Numerical simulations}
To restore the continuous picture one needs to perform a number of bubbles simulations. As we saw
before there are two approaches towards it - either to simulate
the bubble train and then to measure concentration along the pipe, Eq.
\ref{main:mass:transfer:expression}, or to transfer to the reference frame moving with the bulk
velocity $U$ and conduct same measurements. However, both methods do require a tracking of
moving bubbles which is complicated from the numerical point of view. Therefore, one needs to come
up with the simple smaller domain calculations of the mass transfer coefficient, which mimic the
continuous picture with infinite numbers of the bubbles closely.

To avoid complications with moving grids, our
approach is to simulate mass transfer in the reference frame moving with the bubble. Therefore, one
needs to examine the equation for the continuous approach, Eq.
\ref{main:mass:transfer:expression}.

We perform simulations in the frame moving with the bubble (velocity
$\ububble$), that the bubble position stands steady. The bubble velocity $\ububble$ is
different from the bulk velocity $\ugas+\uliq$. In this case we need to perform a coordinate $x$
variable change:
\beqal
\label{theor:average:concentration:time}
&x(t)=\ububble t\\
&C(x)=\cstar \Bigl(1-\exp\bigl(-\vol \frac{x}{\ugas+\uliq}\bigr)\Bigr)\\
&C(t)=\cstar \Bigl(1-\exp\bigl(-\vol\, t \frac{\ububble}{\ugas+\uliq}\bigr)\Bigr).
\feqal
Therefore, one can obtain the volumetric mass transfer coefficient through the concentration:
\beqal
&\vol\, t \frac{\ububble}{\ugas+\uliq}=\ln \frac{\cstar}{\cstar-C(t)}\\
&\volnondim=\frac{\lunit}{\ububble t}\ln \frac{\cstar}{\cstar-C(t)},
\feqal
where the parameter $\vol \frac{\lunit}{\ugas+\uliq}$ is non-dimensional. One can also measure the
volumetric mass transfer coefficient from concentrations given at times $t_1$ and $t_2$:
\beq
\label{theor:continuous:mass:transfer}
\volnondim=\frac{\lunit}{\ububble
(t_2-t_1)}\ln\frac{C^{*}-C(t_1)}{C^{*}-C(t_2)}.
\feq
Expressions (\ref{theor:average:concentration:time}) and (\ref{theor:continuous:mass:transfer}) are
the cornerstones of the present work. The main questions arisen in the present work are to define
$C(t)$ and  the boundary conditions to produce a proper volumetric mass transfer coefficient
$\vol$.

Three possible scenarios of conducting numerical simulations are possible: 
\begin{enumerate}
\item As far as the hydrodynamic flow is periodic, one can make an assumption of the same boundary
conditions applied to the tracer concentration. Therefore, only one unit cell is simulated with
periodic boundary conditions, see Fig. \ref{fig:benchmark}. As well, no tracer leaves a
domain. The situation resembles a situation for the plug flow, where no tracer can leave a unit
domain. We expect such a situation happening for small Capillary numbers $Ca<0.1$, where the film
is thin and saturates fast with the tracer. Though easier to implement, it rises certain critisism
towards the inlet concentration to be equal to the outlet concentration. In the continuous
formulation there is the concentration difference between inlet and outlet involving the volumetric
mass transfer coefficient:
\beq
\volnondim=\ln\frac{\cstar-\cinlet}{\cstar-\coutlet}.
\feq
The concentration $C(t)$ required for calculation the volumetric mass transfer coefficient, Eq.
(\ref{theor:average:concentration:time}), can be taken as the average concentration in the domain:
\beq
C(t)=\frac{\int_{liquid}{C \mathrm{d}V}}{\int{\mathrm{d}V}}.
\end
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/benchmark_middle.eps}\\
\caption{The two-dimensional benchmarks for the  the mass transfer
coefficient (bottom) for the bubble located near the entrance (top) and at the middle of the domain
(bottom). \label{fig:benchmark}}
\end{figure}
\item The second approach is the same as the first one but the concentration is taken as in
\cite{vanbaten-circular}. The characteristic concentration is taken as the inlet/outlet
concentration:
\beqal
\cinlet=\frac{\int{U(y) C(0,y,t) \mathrm{d}y}}{\int{U(0,y) \mathrm{d}y}}\\
\coutlet=\frac{\int{U(y) C(\Lunit,y,t) \mathrm{d}y}}{\int{U(\Lunit,y) \mathrm{d}y}}\\
\cinlet=\coutlet.
\feqal
Therefore the assumptions  of this approach is that the difference between inlet and outlet is not
considerably large and the tracer is well mixed in the slug.

\item
The approach of \citet{vanbaten-circular} is taken. Periodic boundary conditions were used. By the
definition the mass transfer coefficient is
proportional to the gain of the mass in the system divided by the concentration difference
multiplied on the surface area:
\beq
\vol=\frac{\dot{m}}{P \Delta C}\frac{P}{V}=\frac{\dot{m}}{V (\cstar-C(t))},
\feq
where mass difference in the domain can be calculated as:
\beq
\dot{m}=\frac{m_2-m_1}{t_2-t_1}=\frac{\int_{liq}{C(t_2)\mathrm{d}V}-\int_liq{C(t_1)\mathrm{d}V}}{
t_2-t_1 } .
\feq
In the approach of \citet{vanbaten} the inlet and outlet concentrations were taken as the
characteristic concentration $C(t)$.

\item 
The tracer picking up procedure can be introduced
in the manner of the considerations above. We impose the concentration at the entrance to be
constant $C(0,y)=C_{inlet}$, at the outlet we impose the outflow boundary condition $\frac{\partial
C}{\partial x}(L,y)=0$. To be compliant with the considerations above one needs to impose as well
the outflow boundary conditions for the inlet as well. However, without loss of the generality one
can impose $C_{inlet}=0$, since the walls in the reference frame
moving with the bubble are moving from the inlet to the outlet. That means that no tracer
concentration leaves domain from the inlet side, which is true for the capillary numbers larger
than $0.6$.  However, this boundary condition can not be used in the case of small capillary
numbers, since there is a strong circulation in front of the bubble. One can see the sketch of the
approach in Fig. \ref{fig:benchmark:jos}. The longer the numerical domain length the better will be
results. One can estimate the volumetric mass transfer coefficient in this case through the equation
(\ref{small:conc:mass:estimation}) for small concentrations. Note a few considerations:
\begin{itemize}
\item  The considerations above were established for the homogenous velocity field. In the case of
the non-homogenous velocity field one needs to obtain a concentration from the mass flux:
\beq
C_{outlet}=\frac{\int{U(y) C(L_{unit},y) \mathrm{d}y}}{\int{U(y) \mathrm{d}y}}.
\feq
\item Note, that one cannnot use the definition of the mass transfer coefficient as far as it
varies through the bubble due to inhomogenoities of the flow. Therefore, the volumetric mass
transfer coefficient is introduced that takes into the account the bubble geometry.
\end{itemize}
\begin{figure}
\includegraphics[width=\textwidth]{Figures/benchmark_jos.eps}
\caption{The numerical mass transfer benchmark. \label{fig:benchmark:jos}}
\end{figure}

\end{enumerate}

@@

\section{Validation}

\section{Numerical approach}
\label{sec:numerics}
The numerical  procedure of studying mass transfer is presented in the following steps:
\begin{description}
 \item[Flow field] The hydrodynamics for the bubble motion in the periodic domain is obtained
according to the work \cite{kuzmin-binary2d}. The grid resolution is taken to insure grid
independency of results.  
 \item[Bubble reference frame] Once the hydrodynamics is resolved, the mass transfer simulations
are conducted in the moving
reference frame with the bubble. Then the bubble stands still and the flow is coming around the
bubble. We impose a steady concentration of solutant on the surface of the bubble.
\item[Mass transfer] One needs not only perform a Galilean transformation to the reference
frame moving with the bubble but also to
transfer the bubble to the inlet condition. This can insure that the studied phenomenon is a liquid
slug. 
\end{description}

@@

\section{Comparison between different approaches for one unit cell}
We consequently will examine different approaches to the mass transfer definition, Fig.
\ref{fig:comparison:one:unit:cell}:
\begin{description}
\item[I] Definition of \citet{vanbaten-circular} with periodic boundary conditions where a bubble
is located near the entrance of the computational domain, and the characteristic concentration is
the inlet/outlet concentration, Fig \ref{fig:benchmark}. 
\item[II] Definition of \citet{vanbaten-circular} with periodic boundary conditions where a bubble
is located near the entrance of the computational domain, and the characteristic concentration is
the average concentration, Fig \ref{fig:benchmark}.
\item[III] Definition of \citet{vanbaten-circular} with periodic boundary conditions where a
bubble is located in the middle of the computational domain, and the characteristic concentration
is the inlet/outlet concentration.
\item[IV] Definition of \citet{vanbaten-circular} with periodic boundary conditions where a bubble
is located in the middle of the computational domain, and the characteristic concentration is the
average concentration.
\item[V] Definition with the concentration jump between inlet and outlet over one unit cell.
\end{description}
